Supporting Information

### 2018-03-28 SupportingInformation.Rmd
editing, mostly the tests of the nonparametric stats

For a bivariate dataset $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps, 
most implemented using tools from the `VineCopula` package. Code combining these tools is in the
[*OurBiCopSelect*](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [*BIVAN*](https://github.com/sghosh89/BIVAN) repository. 
<!--Shya, links to code are a good idea! Thanks for figuring this out and adding, we should do something like this in other methods sections.-->

@@

The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`. Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}* copulas.
Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC) and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas @burnham2003_modelselection. 
<!--Shya, we here need a reference to the Burnham and Anderson book-->
Descriptions of these copula families can be found in introductory texts on copulas @burnham2003_modelselection, @nelsen2007_copula
<!--Shya to please insert references for both the books on copulas we have-->
and in references cited in `VineCopula`. The lower- and upper-tail dependence of the 
best-fitting member of each family was also provided by `BiCopEst`.

@@

Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas @burnham2003_modelselection
<!--Shya to add reference to Burnham and Anderson, please-->
; likewise for BIC.

@@

<!--Shya, can you please change this figure so it shows u (horizontal) and v (vertical) on the axis labels instead of vi and vj? Often times computer variables do not make good mathematical notation, so I have changed this, I also made changes for similar reasons to other notation elsewhere.-->

Left- and right-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see Sec. \ref{Model_selection} for defintions of $u_i$ and $v_i$)
<!--Shya pls insert ref to the section immediately above this one-->
are related in the left and right tails, respectively, of their 
distributions @nelsen2007_copula. 
<!--Shya, pls cite the Nelson book-->
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Figure \ref{fig_stat_image}). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions.

@@

### Stat-testing of Non-parametric analysis (for limited number of datapoints) \label{Stat_testing}

@@

```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"))}
set.seed(seed)

@@

```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"),mtime("TestStats_multivar_dataset.R"))}
set.seed(seed)
resultsloc<-"./Results/stat_results/stat_testing/"
b<-callfn$paramlist
plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b)
plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b)
plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b)
plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b)
```

>>>>

For a bivariate dataset $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps, 
most implemented using tools from the `VineCopula` package. Code combining these tools is in the
[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository. 
<!--Shya, links to code are a good idea! Thanks for figuring this out and adding, we should do something like this in other methods sections.-->

@@

The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`. Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas @burnham2003_modelselection. 
Descriptions of these copula families can be found in introductory texts on copulas @nelsen2007_copula
<!--Shya to please insert references for the book on copulas by Joe - it is on my desk, come get it please-->
and in references cited in `VineCopula`. The lower- and upper-tail dependence of the 
best-fitting member of each family was also provided by `BiCopEst`.

@@

Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas @burnham2003_modelselection; likewise for BIC.

@@

Left- and right-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see Sec. \ref{Model_selection} for defintions of $u_i$ and $v_i$)
are related in the left and right tails, respectively, of their 
distributions @nelsen2007_copula. 
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Figure \ref{fig_stat_image}). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions.

@@

@@

<!--Shya, why is message=F and warning=F? This is risky unless you ran it already without these and 
know what the warnings are and that we can ignore them. If this is the case, please add a comment here
explaining why it is safe. If you do not have a complete understanding of the warnings and messages,
please do not suppress them. Even if you do understand them, there are probably better ways to handle 
it than supressing them.-->
```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"))}
set.seed(seed)

#Shya, there are a lot of functions in TestStats_multivar_dataset, do we still need all of them? Please delete any that 
#we no longer use! You need to keep code clean!

@@

```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"),mtime("TestStats_multivar_dataset.R"))}

b<-callfn$paramlist
plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b,resultsloc)
plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b,resultsloc)
plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b,resultsloc)
plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b,resultsloc)
```
